\chapter*{Acknowledgment}

	The journey of completing this dissertation has been transformative, made possible by the extraordinary support of
	many remarkable individuals.
	\\
	\\
	I am profoundly grateful to the Port of Auckland, especially Manvi Madan and Valerie Wijaya, whose collaboration
	went far beyond professional courtesy. Their understanding during my health challenges and willingness to adapt to
	my circumstances made this research possible.
	\\
	\\
	Two exceptional mentors guided my journey. To my supervisor, Dr. Kaiqi Zhao, thank you for your intellectual
	guidance, which consistently challenged me to reach higher standards while showing patience and understanding
	when I needed it most. Susanne Cooke, our Industry Engagement Manager, you were more than just a professional guide
	– your timely advice on managing health challenges alongside research responsibilities proved invaluable. Together,
	you shaped my research and understanding of proper academic mentorship.
	\\
	\\
	In the moments when the path seemed most challenging, I found strength through the support of the university's
	student services team. Chanel Tafa from Campus Care and Emma Cooper-Williams, our Student Support Adviser, your
	compassionate guidance helped me navigate personal hurdles while keeping my academic dreams alive. You showed me
	that seeking support is not a sign of weakness but a step toward resilience.
	\\
	\\
	To my beloved wife, your decision to build a new life with me in New Zealand speaks volumes about your love and
	commitment. Your strength was outstanding, holding everything in place as we learned about becoming parents while I
	did this dissertation. Your faith in me has been my guiding light through late nights of research and moments of
	doubt. This achievement is as much yours as it is mine, and sharing this journey of academic growth while preparing
	for parenthood has made it infinitely more meaningful.
	\\
	\\
	This dissertation represents industry research and an example of human connection and support. Each person
	mentioned here has contributed to this scholarly work and my growth as a professional, a husband, and a soon-to-be
	father. For this, I am eternally grateful.