---
output:
  pdf_document: default
  html_document: default
---

# Data Estructure

```{r data-load, cache=TRUE, echo=FALSE}
yard_moves_by_year.df <- read.csv("yard_moves_for_year.csv", header = TRUE)

```

```{r adding-data-colunms, cache=TRUE, echo=FALSE}
yard_moves_by_year.df$year_ <-
  as.integer(substr(yard_moves_by_year.df$period, 1, 4))
yard_moves_by_year.df$month_ <-
  as.integer(substr(yard_moves_by_year.df$period, 6, 7))
yard_moves_by_year.df$day_ <-
  as.integer(substr(yard_moves_by_year.df$period, 9, 10))

yard_moves_by_year.df[yard_moves_by_year.df$mve_gkey == "NA", ]


# Cleaning some data based on MANVI's comments on  24 APR 2024
## ufv_gkey: Feel free to disregard this gkey for your analysis, its an internal key we use to join across tables.
yard_moves_by_year.df <-
  subset(
    yard_moves_by_year.df,
    select = -c(
      che_fetch_login_name,
      che_carry_login_name,
      che_put_login_name
    )
  )

## rehandle_count if NA will be 0
yard_moves_by_year.df$rehandle_count[is.na(yard_moves_by_year.df$rehandle_count)] <-
  0

## twin_fetch and twin_carry if NA will be 0
yard_moves_by_year.df$twin_carry[is.na(yard_moves_by_year.df$twin_carry)] <-
  0
yard_moves_by_year.df$twin_fetch[is.na(yard_moves_by_year.df$twin_fetch)] <-
  0

# Cleaning some data based on MANVI's comments on  26 APR 2024
yard_moves_by_year.df$dist_start[is.na(yard_moves_by_year.df$dist_start)] <-
  0
yard_moves_by_year.df$dist_carry[is.na(yard_moves_by_year.df$dist_carry)] <-
  0
yard_moves_by_year.df$twin_put[is.na(yard_moves_by_year.df$twin_put)] <-
  0


cols_with_na <-
  sapply(yard_moves_by_year.df, function(x)
    any(is.na(x)))
columns_with_na <- names(cols_with_na[cols_with_na])

```

```{r, echo=FALSE}
str(yard_moves_by_year.df)
```


# Years and months Provided

```{r, echo=FALSE, cache=TRUE}
table(yard_moves_by_year.df$year_,yard_moves_by_year.df$month_)
```
# Categorical Data

| ID's Column Name      |   Description                                                                                                          |
|-----------------------|------------------------------------------------------------------------------------------------------------------------|
| nominal_length        |   Container size that could be 20, 40, 45, and 10                                                                      |
| arrive_pos_loctype    |   Type of location arrival                                                                                             |
| last_pos_loctype      |   Last position's type where the container was fetched from, elated to fm_pos_locid                                    |



### category

| Name Category  | Description                                                                    |
|----------------|--------------------------------------------------------------------------------|
| IMPRT          | Import                                                                         |
| EXPRT          | Export                                                                         |
| STRGE          | Storage                                                                        |
| TRSHP          | When a box comes off a vessel and goes on another vessel                       |
| THRGH          | When a box leaves on the same vessel it came on, it just went through          |

### move_kind

| Name Category  | Description                                            |
|----------------|--------------------------------------------------------|
| DLVR           | Delivery                                               |
| LOAD           | Load                                                   |
| YARD           | Yard Move                                              |
| SHFT           | Yard Shift                                             |
| RECV           | Receival                                               |
| RLOD           | Rail load                                              |
| OTHR           | Other                                                  |
| RDSC           | Rail Discharge                                         |
| DSCH           | Discharge                                              |
| SHOB           | Shift O.B                                              |

### freight_kind
| Name Category  | Description                                            |
|----------------|--------------------------------------------------------|
| MTY            | Empty                                                  |
| FCL            | Full Container Load                                    |


## Binary Variables

| ID's Column Name |   Description                                                                                                                           |
|------------------|-----------------------------------------------------------------------------------------------------------------------------------------|
| requires_power   |   Container deigned to be transported as temperature sensitive cargo and the temperatures associated with them are precisely controlled |
| twin_fetch       |   Pairs of containers that were able be twin fetched                                                                                    |
| twin_carry       |   Pairs of containers that can able be twin carried                                                                                     |
| twin_put         |   time when the whole chain of events related to a container_visit_gkey is completed **(TO DOUBLE CHECK)**                              |

## Dates

| ID's Column Name    | Description.                                                                                                                                     |
|---------------------|--------------------------------------------------------------------------------------------------------------------------------------------------|
| t_put               | It is the time completed                                                                                                                         |
| t_dispatch          | This is the time when a container is dispatched somewhere like on a vessel or on truck                                                           |
| t_carry_dispatch    | **Unsure about this one, need to be double checked**                                                                                             |
| t_fetch             | Datetime when the container was fetched                                                                                                          |
| t_discharge         | Datetime when the container is left in a certain position after being fetched (which could be twin_fetched)                                      |
| time_in             | Datetime when the container arrives at the port.  In case of stranships this field could also indicate change of category from storage to export |
| time_out            | It is associated with each stage, if a container was tranship and it changed from import to storage to export, each stage would have a time_out. |

## Numerical Data

| Name Category           | Description                                                                                                             |
|-------------------------|-------------------------------------------------------------------------------------------------------------------------|
| rehandle_count          | How many times a container was rehandled. NOT needless                                                                  |
| goods_and_ctr_wt_kg     | Total weight goods and container (Container weight and Goods weight)                                                    |
| dist_start              | Distance where the container started to ve moved, related to dist_carry (Need to be verified)                           |
| dist_carry              | Distance the container was carried during the move would be correlated to distance it started at (Need to be verified)  |


## Identifiers

| ID's Column Name  | Description                                                                                        |
|-------------------|----------------------------------------------------------------------------------------------------|
| mve_gkey          | Unique identifier for each move                                                                    |
| fm_pos_locid      | FROM Position location id associated with fm_pos_name                                              |
| fm_pos_name       | Name of the location associated with fm_pos_locid                                                  |
| to_pos_locid      | TO Position location associated with to_pos_name                                                   |
| to_pos_name       | Name of the location associated with to_pos_locid                                                  |
| inv_unit_id       | Container ID, same even if revisit the port                                                        |
| bizunit_id        | Line operator's ID.                                                                                |
